---
title: 'Add a GeoJSON/sf source'
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```


## Preparing the data

A GeoJSON source is a format used to represent geographical features (as well as their non-geographic attributes). These sources can hold features like polygons, lines, points, circles, and more.
In R a GeoJSON object will contain the `geojson` class.

An `sf` object is a [simple feature](https://r-spatial.github.io/sf/articles/sf1.html) data frame that contains a `geometry` column along with other columns containing a variety of attributes.
In R an `sf` object will contain the `sf` class.

For this example, we use the [quakes](https://www.rdocumentation.org/packages/datasets/versions/3.6.2/topics/quakes) dataset that comes with base R.
`quakes` is, by default, just a data frame. We use the `sf` package to turn it into a simple feature data frame.

```{r prepare-data}
library(sf)
library(geojsonsf)
library(dplyr)
library(toro)

data(quakes)
quakes_data <- quakes |>
  dplyr::mutate(hover = paste0("Magnitude: ", mag, "</br>Depth: ", depth)) |>
  sf::st_as_sf(coords = c("long", "lat"), crs = 4326)
```

At this point the data is still not GeoJSON data (it is an `sf` object), however, the `add_source()` function contains the ability to turn an `sf` object into a `geojson` object.
This means that we can pass `quakes_data` to `add_source()` and it will be internally converted from `sf` to `geojson`.

For the purpose of this example, we will split `quakes_data` into two objects:

1. An `sf` object containing quakes with a magnitude below 6
2. A `geojson` object containing quakes with a magnitude of 6 or above

```{r geojson-transform}
# Keep as a sf object
sf_quakes_data <- quakes_data |>
  dplyr::filter(mag < 6)
print(sf_quakes_data)

# Transform into a geojson object
geojson_quakes_data <- quakes_data |>
  dplyr::filter(mag >= 6) |>
  geojsonsf::sf_geojson()
print(geojson_quakes_data)
```


## Adding a GeoJSON/sf source

There are two ways to add a GeoJSON/sf source to your map:

### The add_source() function

Using the `add_source()` function, you can add a GeoJSON source by passing an `sf` or `geojson` object directly to the `data` argument.
`sf` objects will be automatically converted to GeoJSON format when added as a source.
You can then use this source when adding layers to your map by specifying the `source` argument in the layer functions.

You can pass either a `map`or `mapProxy` object to `add_source()`.

```{r add-source-map}
map() |>
  set_bounds(bounds = quakes_data) |>
  add_source("sf_quakes_data", data = sf_quakes_data) |>
  add_source("geojson_quakes_data", data = geojson_quakes_data) |>
  add_circle_layer(
    id = "quakes_circles_sf",
    source = "sf_quakes_data",
    hover_column = "hover",
    paint = get_paint_options("circle", list(colour = "orange"))
  ) |>
    add_circle_layer(
    id = "quakes_circles_geojson",
    source = "geojson_quakes_data",
    hover_column = "hover",
    paint = get_paint_options("circle", list(colour = "red"))
  )
```


### Adding the source directly in the layer function

Alternatively, you can add a GeoJSON/sf source directly within a layer function by passing a sf object or GeoJSON string to the `source` argument.
In this case the source ID will be automatically generated by appending the layer ID to `source-`.
For example, if you add a layer with `id = "my_layer"` and `source` is a sf object, the source will be added to the map with the ID `"source-my_layer"`.
This allows you to add the source and layer in a single step, without needing to call `add_source()` separately.


```{r advanced-map}
map() |>
  set_bounds(bounds = quakes_data) |>
  add_circle_layer(
    id = "quakes_circles_sf",
    source = sf_quakes_data,
    hover_column = "hover",
    paint = get_paint_options("circle", list(colour = "orange"))
  ) |>
    add_circle_layer(
    id = "quakes_circles_geojson",
    source = geojson_quakes_data,
    hover_column = "hover",
    paint = get_paint_options("circle", list(colour = "red"))
  )
```
