---
title: "Getting started with toro"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{getting-started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
format: html
editor: visual
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(toro)
```

## Installation

To install the development version from Github:

```r
remotes::install_github("Epi-interactive-Ltd/toro")
```

Once installed, this package can be used in the R console and within [Shiny](https://shiny.posit.co/) applications.

## Basic Usage

### Default Map

To create the default map widget just call the `map` function.

```{r basic-usage}
map()
```

### Customise With Map Options

By passing parameters to the `map` function you can customise several features of the map.

```{r customise-default}
map(
  style = "natgeo",
  center = c(174, -41),
  zoom = 2
)
```

### Adding Layers

You can add layers to the map using layer functions like `add_symbol_layer`, `add_fill_layer`, etc.
For more information about using layers see the [layers vignette](https://epi-interactive-ltd.github.io/toro/articles/layers.html).

Data layers like fill, marker, and line layers need source data for the layers to show anything.
There is two ways to provide this source data to a layer.

#### 1. Passing the data inside the layer function

```{r source-in-function}
map()
```

#### 2. Adding the source data before adding the layer

```{r source-before-function}
map()
```
