---
title: "map-tiles"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{map-tiles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(toro)
```

## Map Tiles


You can find the tile options by using `get_tile_options()`.


```{r tile-options, results='asis'}
get_tile_options()
```

When declaring a map, you can specify the style and any tiles you want the map to start with
by using the `style` parameter.

By default, the `style` is set to `light-grey`, but you can change it to any of the options
available.

```{r basic-tiles, results='asis'}
map(style="satellite")
```

If you want the ability for the map to change tiles you need to set the `loadedTiles` parameter to
a vector of the tiles you want to be able to switch between.

```{r loaded-tiles, results='asis'}
map(
  style="satellite",
  loadedTiles = c("satellite", "light-grey", "natgeo")
)
```

## Changing Tiles

### set_tile_layer()

You can use the `set_tile_layer()` function to update the tiles on the map to any of the options
defined in `loadedTiles` when the map was created.

```{r set-tiles, results='asis'}
library(shiny)
library(toro)


ui <- fluidPage(
  toro::mapOutput("map"),
  selectInput(
    "tiles",
    "Tiles",
    choices = toro::get_tile_options(),
    selected = "satellite"
  ),
)

server <- function(input, output, session) {
  output$map <- toro::renderMap({
    toro::map(style = "satellite", loadedTiles = toro::get_tile_options())
  })

  observe({
    req(input$tiles)
    toro::mapProxy("map") |>
      set_tile_layer(input$tiles)
  })
}

if (interactive()) {
  shinyApp(ui, server)
}
```

### Tile Controllers

You can also change a maps tiles via a tile controller.
A tile controller must be placed inside a control panel.


For more information on adding controls to a map see the [controls vignette](https://epi-interactive-ltd.github.io/toro/articles/controls.html).


```{r tile-controller, results='asis'}
library(shiny)
library(toro)

all_tiles <- toro::get_tile_options()

ui <- fluidPage(
  toro::mapOutput("map")
)

server <- function(input, output, session) {
  output$map <- toro::renderMap({
    toro::map(style = "satellite", loadedTiles = all_tiles) |>
      add_control_panel(
        panel_id = "demo-panel"
      ) |>
      add_tile_selector_control(
        available_tiles = all_tiles,
        panel_id = "demo-panel"
      )
  })

}

if (interactive()) {
  shinyApp(ui, server)
}
```

