---
title: 'Add a feature service source'
output: rmarkdown::html_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Resources

- [ArcGIS Feature Service documentation](https://developers.arcgis.com/rest/services-reference/enterprise/feature-service/)

---

## What is a feature service source?

A Feature Service is a web service that provides access to geographic features and their attributes. In toro, you can add a feature service as a data source for your map layers using the `add_feature_server_source` function.


## Adding a feature service source

### Basic example

Most of the time, you can simply provide the URL of the feature service and toro will handle the rest. The default query parameters will be appended to the URL to retrieve the data in GeoJSON format.

```{r base-map}
library(toro)

map() |>
  add_feature_server_source(
    "https://services1.arcgis.com/VwarAUbcaX64Jhub/arcgis/rest/services/World_Exclusive_Economic_Zones_Boundaries/FeatureServer",
    "eez"
  ) |>
  add_line_layer(id = "eez_lines", source = "eez")
```


### Advanced example

If you need more control over the query parameters, you can provide the full query URL directly in the `source_url` argument and set `append_query_url` to an empty string to prevent appending the default query parameters.

```{r advanced-map}
library(toro)

map() |>
  add_feature_server_source(
    "https://services8.arcgis.com/AYGZtmUtpARUKBlB/arcgis/rest/services/Te_Reo_M%C4%81ori_Place_Names/FeatureServer/4/query?where=1=1&outFields=*&f=geojson",
    "maori_awa_data",
    append_query_url = "" # We supply the full query URL here, so we set this to an empty string to avoid appending the default query parameters.
  ) |>
  add_line_layer(
    id = "awa_lines",
    source = "maori_awa_data",
    hover_column = "name_mi",
    popup_column = "name"
  )
```
