---
title: "Animations"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/"
)
```

toro provides the ability to animate certain aspects of the map.

## Setup

For the following examples we'll use a line representing a route around Aotearoa.

```{r setup}
library(sf)
library(toro)

nz_line <- sf::st_sf(
  id = 1,
  geometry = sf::st_sfc(
    sf::st_linestring(
      cbind(
        c(
          172.2041,
          182.2494,
          168.1286,
          163.9383,
          171.7449
        ),
        c(
          -32.56960,
          -38.85522,
          -48.49620,
          -46.43999,
          -34.62533
        )
      )
    ),
    crs = 4326
  )
)
```


## Basic example

For the most basic example, add a route (line) and then use `add_animation_controls()` to
add controls to the map for animating along the route.

```{r basic-animation}
map() |>
  add_route(route_id = "nz_route", points = nz_line) |>
  add_animation_controls(route_id = "nz_route")
```


## Advanced examples

### Speed control

You can add a customisable speed control.
By default, there are 3 speed options: 0.5x, 1x, and 2x.
You can change the speed options by passing custom values and labels to the `add_animation_controls()` function.

```{r speed-animation}
map() |>
  add_route(route_id = "nz_route", points = nz_line) |>
  add_animation_controls(
    route_id = "nz_route",
    include_speed_control = TRUE,
    speed_values = c(0.3, 0.5, 1, 2, 4),
    speed_labels = c("Slower", "Slow", "Default", "Fast", "Faster")
  )
```

### Animating icon

You can also supply an icon to be animated along the route.
By default, toro has a built in image with the ID "toro-pin" that you can use for this purpose, or you can add your own custom image.
To use a custom image, use `add_image()` to add the image and then pass the `settings.animatingIcon` parameter to `add_route()` to specify the icon you want to animate along the route.


```{r icon-animation}
map() |>
  add_route(
    route_id = "nz_route",
    points = nz_line,
    settings = c(list(
      animatingIcon = get_layout_options(
        "symbol",
        list(icon_image = "toro-pin", icon_size = 0.5)
      )
    ))
  ) |>
  add_animation_controls(route_id = "nz_route")
```

### Other customisations

You can set `dropVisited` to `TRUE` in the animation settings to leave a marker on the map at each point that has been visited during the animation.

You can also change the styles of the different parts of the route using the `animatedRouteStyle`, `unanimatedRouteStyle`, and `visitedRouteStyle` settings.


```{r customisations-animation}
map() |>
  add_route(
    route_id = "nz_route",
    points = nz_line,
    settings = c(list(
      dropVisited = TRUE,
      routeLine = list(
        "line-color" = "#11a366",
        "line-width" = 2,
        "line-dasharray" = c(2, 4)
      ),
      visitedPoints = list(
        "circle-color" = "#20de60",
        "circle-radius" = 4
      ),
      animatingPoint = list(
        "circle-color" = "#dbe811",
        "circle-radius" = 8
      )
    ))
  ) |>
  add_animation_controls(route_id = "nz_route")
```



